---
title: "Analysing the factors affecting income of data scientists"
author: "Abhishek"
date: "1/2/2020"
output:
  word_document: default
  html_document: default
---
```{r}
```


```{r}
#Importing libraries
#library(tidyverse)
library(gridExtra)
library(scales)
#library(plotly)
library(RColorBrewer)
library("dplyr")
library("ggplot2")
#library(skimr)
library(plyr)
library(dplyr)
library(stringr)
library(forcats)
#library(descr)
#library(repr)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#*************************************INTRODUCTION*******************************************#
This project aims to explore the factors that affect salary and income of data scientists across the globe. The dataset used for this project is taken from kaggle. It is based on a survey conducted in 2018 about personal, professional and technological details of data scientists across the globe. The survey consisted of 50 multiple choice questions. All the data in the dataset is categorical. More details are given in the word document.

#*********************************FRAMEWORK FOR ANALYSIS***************************************#
For analysing the data and creating the visualisations, I focussed on three main types of parameters: Socio-Demographic, Employment Related and Technological. For each section, I chose variables that affected the income variable in a practical sort of way, i.e. those which can be worked upon, inorder to present the audience with parameters that they can improve or work on. More details are given in the word document.

#*********************************VARIABLE SELECTION***************************************#
I selected the variables based on the criteria I have mentioned above. As all the variables were categorical, it was not possible to calculate a heatmap to get the correlation coefficient. So I chose the variables on the basis of my domain knowledge of the IT industry. Some variables were refactored as required. More details are given in the word document.
```{r }
#Reading the file
df <- read.csv('D:\\MS\\R\\Project\\multipleChoiceResponses.csv')
```



```{r}
#Gender
options(repr.plot.width=5, repr.plot.height=4)
gender_plot <- df %>%
filter(Q1=="Male"|Q1=="Female")%>% 
ggplot(aes(Q1))+geom_bar(aes(fill=Q1))+scale_fill_brewer(palette="Set2",  na.value = "gray")+xlab("Gender")+theme(legend.position="none")
ggsave("gender.png")
gender_plot
```





```{r }

```





```{r }
#Refactoring income to 4 income levels
df <- mutate(df,Q9=ifelse(Q9=="0-10,000"|Q9=="10-20,000"|Q9=="20-30,000"|Q9=="30-40,000"|Q9=="40-50,000","Low",
                              ifelse(Q9=="50-60,000"|Q9=="60-70,000"|Q9=="70-80,000"|Q9=="80-90,000"|Q9=="90-100,000","Medium",
                                     ifelse(Q9=="What is your current yearly compensation (approximate $USD)?","What is your current yearly compensation (approximate $USD)?",
                                            ifelse(Q9==""|Q9=="I do not wish to disclose my approximate yearly compensation","NA","High")))))
  
```

```{r }
#Income plot
income_plot<-df[-1,]%>%
  filter(!Q9=="I do not wish to disclose my approximate yearly compensation"&!is.na(Q9))%>%
  ggplot(aes(Q9))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Income Levels", fill = "Income")
ggsave("income_plot.png")
income_plot

```

```{r }
#Refactoring names of countries
df$Q3 <- mapvalues(df$Q3, from = c("Iran, Islamic Republic of..."), to  = c("Iran"))
df$Q3 <- mapvalues(df$Q3, from = c("United Kingdom of Great Britain and Northern Ireland"), to  = c("UK"))
df$Q3 <- mapvalues(df$Q3, from = c("United States of America"), to  = c("USA"))
df$Q3 <- mapvalues(df$Q3, from = c("Republic of Korea"), to  = c("South Korea"))
df$Q3 <- mapvalues(df$Q3, from = c("Hong Kong (S.A.R.)"), to  = c("Hong Kong"))
```


```{r }
#Income by country
income_country<-df[-1,]%>%
  filter(Q3!=""& Q3!="I do not wish to disclose my location")%>%
  ggplot(aes(Q3))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Country of Residence", fill = "Income")
ggsave("income_country.png")
income_country
```


```{r }
#High income level by country
income_country_high<-df[-1,]%>%
  filter(Q3!=""& Q3!="I do not wish to disclose my location"&Q9 == "High")%>%
  ggplot(aes(Q3))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Country of Residence", fill = "Income")
ggsave("income_country_high.png")
income_country_high
```


```{r }
#High income level by country - pie-chart
high_income_pie_chart<-df[-1,]%>%
  filter(Q3!=""& Q3!="I do not wish to disclose my location"&Q9 == "High")%>%
  ggplot(aes(x="", y=Q3, fill=Q3)) + geom_bar(stat="identity", width=1)+
  coord_polar("y", start=0)+labs(x = NULL, y = NULL, fill = NULL, title = "High Income-Country Wise")+theme_classic()   + theme(axis.line = element_blank(),
          axis.text = element_blank(),
          axis.ticks = element_blank(),
          plot.title = element_text(hjust = 0.5, color = "#666666"))
ggsave("high_income_pie_chart.png")
high_income_pie_chart
```


```{r }
#Age group
age_group<-df[-1,]%>%
  filter(!is.na(Q2))%>%
  ggplot(aes(Q2))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Age Group", fill = "Income")
ggsave("age_group.png")
age_group
```

```{r}
#Refactoring education levels
df$Q4 <- mapvalues(df$Q4, from = c("Some college/university study without earning a bachelor’s degree"), to  = c("Non University"))
df$Q4 <- mapvalues(df$Q4, from = c("No formal education past high school"), to  = c("Non University"))
df$Q4 <- mapvalues(df$Q4, from = c("Bachelor’s degree"), to  = c("Bachelors"))
df$Q4 <- mapvalues(df$Q4, from = c("Master’s degree"), to  = c("Masters"))
df$Q4 <- mapvalues(df$Q4, from = c("Doctoral degree"), to  = c("Doctoral"))

```


```{r}
#Education level
education<-df[-1,]%>%
  filter(Q4!="")%>%
  ggplot(aes(Q4))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Education Levels", fill = "Income")
ggsave("education.png")
education
```


```{r}
#Refactoring education stream
df$Q5 <- mapvalues(df$Q5, from = c("A business discipline (accounting, economics, finance, etc.)"), to  = c("Business"))
df$Q5 <- mapvalues(df$Q5, from = c("Computer science (software engineering, etc.)"), to  = c("Comp.Sc"))
df$Q5 <- mapvalues(df$Q5, from = c("Engineering (non-computer focused)"), to  = c("Engineering"))
df$Q5 <- mapvalues(df$Q5, from = c("Environmental science or geology"), to  = c("Geology"))
df$Q5 <- mapvalues(df$Q5, from = c("Fine arts or performing arts"), to  = c("Arts"))
df$Q5 <- mapvalues(df$Q5, from = c("Humanities (history, literature, philosophy, etc.)"), to  = c("Humanities"))
df$Q5 <- mapvalues(df$Q5, from = c("Information technology, networking, or system administration"), to  = c("Comp.Sc"))
df$Q5 <- mapvalues(df$Q5, from = c("Mathematics or statistics"), to  = c("Maths"))
df$Q5 <- mapvalues(df$Q5, from = c("Medical or life sciences (biology, chemistry, medicine, etc.)"), to  = c("Life.Sc"))
df$Q5 <- mapvalues(df$Q5, from = c("Physics or astronomy"), to  = c("Physics"))
df$Q5 <- mapvalues(df$Q5, from = c("Social sciences (anthropology, psychology, sociology, etc.)"), to  = c("Humanities"))
```


```{r}
#Education stream
ed_stream<-df[-1,]%>%
  filter(Q5!="")%>%
  ggplot(aes(Q5))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Stream of Education", fill = "Income")
ggsave("ed_stream.png")
ed_stream
```


```{r}
#Education stream facet wrap
ed_stream_facet<-df[-1,]%>%
  filter(Q5!="")%>%
  ggplot(aes(Q5))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+facet_wrap(~Q9)+labs(x = "Stream of Education", fill = "Income")
ggsave("ed_stream_facet.png")
ed_stream_facet
```


```{r}
#Job Title
job<-df[-1,]%>%
  filter(Q6!="")%>%
  ggplot(aes(Q6))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Job Title", fill = "Income")
ggsave("job.png")
job
```
```{r}
#Programming language
p_lang<-df[-1,]%>%
  filter(Q17!="")%>%
  ggplot(aes(Q17))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Programming Language", fill = "Income")
ggsave("p_lang.png")
p_lang
```


```{r}
#Experience in machine learning
age<-df[-1,]%>%
  filter(!is.na(Q25)&Q25!="I have never studied machine learning but plan to learn in the future"&Q25!="I have never studied machine learning and I do not plan to")%>%
  ggplot(aes(Q25))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Age", fill = "Income")
ggsave("age.png")
age
```


```{r}
#Creating dataframe with only high income data
df_high <- df[df$Q9=="High",]
```



```{r}
#Languages which they use
p_lang_high<-df_high[-1,]%>%
  filter(Q17!="")%>%
  ggplot(aes(Q17))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Programming Language", fill = "Income")
ggsave("p_lang_high.png")
p_lang_high
```

```{r}
#People who work in Python
summary(df_high$Q17=="Python")
```

```{r}
#Calculating percentage of people who work in Python
(888/(888+1330))*100

```


```{r}
#Data Types which they work upon
dtype_high<-df_high[-1,]%>%
  filter(Q32!="")%>%
  ggplot(aes(Q32))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Data Type", fill = "Income")
ggsave("dtype_high.png")
dtype_high
```


```{r}
#ML Frameworks they use most
ml_high<-df_high[-1,]%>%
  filter(Q20!="")%>%
  ggplot(aes(Q20))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Machine Learning Framework", fill = "Income")
ggsave("ml_high.png")
ml_high
```
```{r}
#Visualisation Tools they use most
vistools<-df_high[-1,]%>%
  filter(Q22!="")%>%
  ggplot(aes(Q22))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Visualisation Tools", fill = "Income")
ggsave("vistools.png")
vistools
```
```{r}
#Industry they work for
industry<-df_high[-1,]%>%
  filter(Q7!="")%>%
  ggplot(aes(Q7))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Industry", fill = "Income")
ggsave("industry.png")
industry
```




```{r}
da_tools<-df_high[-1,]%>%
  filter(Q12_MULTIPLE_CHOICE!="")%>%
  ggplot(aes(Q12_MULTIPLE_CHOICE))+geom_bar(aes(fill=Q9))+coord_flip()+labs(x = "Data Analysis Tool", fill = "Income")
ggsave("da_tools.png")
da_tools
```
```{r}
olr<-df_high[-1,]%>%
  filter(Q37!="")%>%
  ggplot(aes(Q37))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Online Learning Resources", fill = "Income")
ggsave("olr.png")
olr
```


```{r}
#Languages which they recommend
pl_high<-df_high[-1,]%>%
  filter(Q18!="")%>%
  ggplot(aes(Q18))+geom_bar(aes(fill=Q9))+theme(axis.text.x = element_text(angle = 90))+labs(x = "Programming Language Suggested", fill = "Income")
ggsave("pl_high.png")
pl_high
```

```{r}
#People who recommend Python
summary(df_high$Q18=="Python")
```


```{r}
#Calculating percentage of people who recommend Python
(1507/(1507+711))*100
```

